---
title: "ImportFromSoftwareSta"
author: "Renato Erazo"
date: "8/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## haven

```{r}
install.packages("haven")
library(haven)
```


```{r}
library(haven)
baseDatos <- read_sav(file.path("./01ImportData","datasets","vivienda.sav"))
```


Transformar los datos de opciones de spss a datos de texto en lascolumnas

```{r}
baseDatos$Provincia <- as.character(as_factor(baseDatos$Provincia))
baseDatos$Regional <- as.character(as_factor(baseDatos$Regional))
baseDatos$Área <- as.character(as_factor(baseDatos$Área))
baseDatos$Vivienda <- as.character(as_factor(baseDatos$Vivienda))
baseDatos$VI02 <- as.character(as_factor(baseDatos$VI02))
baseDatos$VI03 <- as.character(as_factor(baseDatos$VI03))
baseDatos$VI04 <- as.character(as_factor(baseDatos$VI04))

```

Exercise
Import STATA data with haven
Next up are STATA data files; you can use read_dta() for these.

When inspecting the result of the read_dta() call, you will notice that one column will be imported as a labelled vector, an R equivalent for the common data structure in other statistical environments. In order to effectively continue working on the data in R, it's best to change this data into a standard R class. To convert a variable of the class labelled to a factor, you'll need haven's as_factor() function.

In this exercise, you will work with data on yearly import and export numbers of sugar, both in USD and in weight. The data can be found at: http://assets.datacamp.com/production/course_1478/datasets/trade.dta

Instructions
100 XP
Import the data file directly from the URL using read_dta(), and store it as sugar.
Print out the structure of sugar. The Date column has class labelled.
Convert the values in the Date column of sugar to dates, using as.Date(as_factor(___)).
Print out the structure of sugar once more. Looks better now?


```{r}
# haven is already loaded
library(haven)
# Import the data from the URL: sugar

sugar <- read_dta("http://assets.datacamp.com/production/course_1478/datasets/trade.dta")
# Structure of sugar
str(sugar)

# Convert values in Date column to dates
sugar$Date <- as.Date(as_factor(sugar$Date))


# Structure of sugar again
str(sugar)
```


Exercise
Import SPSS data with haven
The haven package can also import data files from SPSS. Again, importing the data is pretty straightforward. Depending on the SPSS data file you're working with, you'll need either read_sav() - for .sav files - or read_por() - for .por files.

In this exercise, you will work with data on four of the Big Five personality traits for 434 persons (Source: University of Bath). The Big Five is a psychological concept including, originally, five dimensions of personality to classify human personality. The SPSS dataset is called person.sav and is available in your working directory.

Instructions
100 XP
Instructions
100 XP
Use read_sav() to import the SPSS data in "person.sav". Name the imported data frame traits.
traits contains several missing values, or NAs. Run summary() on it to find out how many NAs are contained in each variable.
Print out a subset of those individuals that scored high on Extroversion and on Agreeableness, i.e. scoring higher than 40 on each of these two categories. You can use subset() for this.


```{r}
# haven is already loaded
library(haven)
# Import person.sav: traits
traits <- read_sav("D:/Principal/2020/Cursos/DataScience/Import-Cleaning-DataCamp/01ImportData/person.sav")

# Summarize traits
summary(traits)

# Print out a subset
subset(traits,Extroversion > 40 & Agreeableness > 40)
```

Exercise
Factorize, round two

In the last exercise you learned how to import a data file using the command read_sav(). With SPSS data files, it can also happen that some of the variables you import have the labelled class. This is done to keep all the labelling information that was originally present in the .sav and .por files. It's advised to coerce (or change) these variables to factors or other standard R classes.

The data for this exercise involves information on employees and their demographic and economic attributes (Source: QRiE). The data can be found on the following URL:

http://s3.amazonaws.com/assets.datacamp.com/production/course_1478/datasets/employee.sav
Instructions
100 XP

    Import the SPSS data straight from the URL and store the resulting data frame as work.
    Display the summary of the GENDER column of work. This information doesn't give you a lot of useful information, right?
    Convert the GENDER column in work to a factor, the class to denote categorical variables in R. Use as_factor().
    Once again display the summary of the GENDER column. This time, the printout makes much more sense.



```{r}
# haven is already loaded

# Import SPSS data from the URL: work
work <- read_sav("http://s3.amazonaws.com/assets.datacamp.com/production/course_1478/datasets/employee.sav")

# Display summary of work$GENDER
summary(work$GENDER)


# Convert work$GENDER to a factor

work$GENDER <- as_factor(work$GENDER)

# Display summary of work$GENDER again
summary(work$GENDER)
```

## ======================================= foreign package core R

```{r}
install.packages("foreign")
library(foreign)
```


Exercise
Import STATA data with foreign (1)

The foreign package offers a simple function to import and read STATA data: read.dta().

In this exercise, you will import data on the US presidential elections in the year 2000. The data in florida.dta contains the total numbers of votes for each of the four candidates as well as the total number of votes per election area in the state of Florida (Source: Florida Department of State). The file is available in your working directory, you can download it here if you want to experiment some more.
Instructions
100 XP

    Load the foreign package; it's already installed on DataCamp's servers.
    Import the data on the elections in Florida, "florida.dta", and name the resulting data frame florida. Use read.dta() without specifying extra arguments.
    Check out the last 6 observations of florida with tail()

```{r}
# Load the foreign package
library(foreign)

# Import florida.dta and name the resulting data frame florida
florida <- read.dta("florida.dta")

# Check tail() of florida
tail(florida)
```

Exercise
Import STATA data with foreign (2)

Data can be very diverse, going from character vectors to categorical variables, dates and more. It's in these cases that the additional arguments of read.dta() will come in handy.

The arguments you will use most often are convert.dates, convert.factors, missing.type and convert.underscore. Their meaning is pretty straightforward, as Filip explained in the video. It's all about correctly converting STATA data to standard R data structures. Type ?read.dta to find out about about the default values.

The dataset for this exercise contains socio-economic measures and access to education for different individuals (Source: World Bank). This data is available as edequality.dta, which is located in the worldbank folder in your working directory.
Instructions
100 XP

    Specify the path to the file using file.path(). Call it path. Remember the "edequality.dta" file is located in the "worldbank" folder.
    Use the path variable to import the data file in three different ways; each time show its structure with str():
    edu_equal_1: By passing only the file path to read.dta().
    edu_equal_2: By passing the file path, and setting convert.factors to FALSE.
    edu_equal_3: By passing the file path, and setting convert.underscore to TRUE.

```{r}
# foreign is already loaded

# Specify the file path using file.path(): path
path <- file.path("worldbank","edequality.dta")

# Create and print structure of edu_equal_1
edu_equal_1 <- read.dta(path)
str(edu_equal_1)

# Create and print structure of edu_equal_2
edu_equal_2 <- read.dta(path,convert.factors = FALSE)
str(edu_equal_2)


# Create and print structure of edu_equal_3
edu_equal_3 <- read.dta(path,convert.underscore = TRUE)
str(edu_equal_3)

```

How many observations/individuals of Bulgarian ethnicity have an income above 1000?

```{r}
nrow(subset(edu_equal_1,income > 1000 & ethnicity_head == "Bulgaria"))
```

Exercise
Import SPSS data with foreign (1)

All great things come in pairs. Where foreign provided read.dta() to read Stata data, there's also read.spss() to read SPSS data files. To get a data frame, make sure to set to.data.frame = TRUE inside read.spss().

In this exercise, you'll be working with socio-economic variables from different countries (Source: Quantative Data Analysis in Education). The SPSS data is in a file called international.sav, which is in your working directory. You can also download it here if you want to play around with it some more.
Instructions
100 XP

    Import the data file "international.sav" and have R convert it to a data frame. Store this data frame as demo.
    Create a boxplot of the gdp variable of demo.

```{r}
# foreign is already loaded

# Import international.sav as a data frame: demo
demo <- read.spss("international.sav",to.data.frame = TRUE)

# Create boxplot of gdp variable of demo
boxplot(demo$gdp)

```

Excursion: Correlation

If you're familiar with statistics, you'll have heard about Pearson's Correlation. It is a measurement to evaluate the linear dependency between two variables, say
and . It can range from -1 to 1; if it's close to 1 it means that there is a strong positive association between the variables. If is high, also tends to be high. If it's close to -1, there is a strong negative association: If is high, tends to be low. When the Pearson correlation between two variables is 0, these variables are possibly independent: there is no association between and

.

You can calculate the correlation between two vectors with the cor() function. Take this code for example, that computes the correlation between the columns height and width of a fictional data frame size:

cor(size$height, size$width)

The data you've worked with in the previous exercise, international.sav, is again available in your working directory. It's now up to import it and undertake the correct calculations to answer the following question:

What is the correlation coefficient for the two numerical variables gdp and f_illit (female illiteracy rate)?
Instructions
50 XP
Possible Answers

    The correlation is very close to 0. Therefore, no association is existing between female illiteracy and GDP for the data set that is used.
    The correlation is around -0.45. There is a negative correlation, but it is rather weak.
    The correlation is almost equal to +1. GDP and female illiteracy are almost perfectly, positive correlated.
    The correlation is around +0.45. There is a positive correlation, but it is rather weak.
    
    
    
```{r}
cor(demo$gdp,demo$f_illit)
```
    



