---
title: "Working with Data in the Tidyverse"
author: "Renato Erazo"
date: "18/12/2020"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ================== 01 Explore your data

Read a CSV file

In this exercise, you'll use read_csv() twice. The first time you will only specify the filename, but you'll notice a problem with the imported data. The second time you'll use a new argument called skip to fix the problem. Remember to use ?read_csv to read more about arguments like skip and how to use them.

The data you'll work with is from "The Great British Bake-Off." The file "bakeoff.csv" contains data for each episode of the show, organized by series and baker.

This course touches on a lot of concepts you may have forgotten, so if you ever need a quick refresher, download the Tidyverse Cheat Sheet and keep it handy!
Instructions 2/4
25 XP

    2
    3
    4

    Use read_csv() to read in "bakeoff.csv", and assign it to a new dataset bakeoff using the assignment operator (<-).
```{r}
# Load readr
library(readr)

# Create bakeoff from "bakeoff.csv"
bakeoff <- read_csv("bakeoff.csv")
```

Exercise
Read a CSV file

In this exercise, you'll use read_csv() twice. The first time you will only specify the filename, but you'll notice a problem with the imported data. The second time you'll use a new argument called skip to fix the problem. Remember to use ?read_csv to read more about arguments like skip and how to use them.

The data you'll work with is from "The Great British Bake-Off." The file "bakeoff.csv" contains data for each episode of the show, organized by series and baker.

This course touches on a lot of concepts you may have forgotten, so if you ever need a quick refresher, download the Tidyverse Cheat Sheet and keep it handy!
Instructions 4/4
25 XP

    4

    Adapt your code to read in "bakeoff.csv" again, but this time, use the skip argument to skip the first line before reading the data. Print again to view it. Remember you can use ?read_csv
    
```{r}
# Load readr
library(readr)

# Create bakeoff but skip first row
bakeoff <- read_csv("bakeoff.csv", skip = 1)

# Print bakeoff
bakeoff
```
    
 Exercise
Assign missing values

The read_csv() function also has an na argument, which allows you to specify value(s) that represent missing values in your data. The default values for the na argument are c("", "NA"), so both are recoded as missing (NA) in R. When you read in data, you can add additional values like the string "UNKNOWN" to a vector of missing values using the c() function to combine multiple values into a single vector.

The is.na() function is also helpful for identifying rows with missing values for a variable.
Instructions 2/4
25 XP

    2
    3
    4

    Add a filter() line after the pipe (%>%) to show only the rows in the showstopper variable coded as "UNKNOWN".
```{r}
# Load dplyr
library(dplyr)
library(readr)

# Filter rows where showstopper is UNKNOWN
#bakeoff %>% filter(showstopper == "UNKNOWN")
```
   
Import bakers data


```{r}
library(readr)



```
Exercise
Arrange and glimpse

From here, if we don't ask you to load a package, you can assume it's already loaded.

You can combine glimpse() with other functions in a sequence using the pipe (%>%) operator. For example, you can use other dplyr functions like arrange first, then use glimpse by adding a line after the final pipe (%>%):

bakers_mini %>% 
  arrange(age) %>% 
  glimpse() # no argument needed here

Take a glimpse of the bakeoff data we imported in the first set of exercises. On which date did the first episode of the show air in the US?

(Pressing Enter in the console will not add a new line, but will run the code on the current line. If you'd like to add a new line before running your code, you can use Shift+Enter)

```{r}

```

#=========================== 02 Tame your data

Exercise
Cast a column to a date

As we saw in the video, a good workflow for parsing dates using readr is to, for example:

    Use parse_date("2012-14-08", format = "%Y-%d-%m") first, then
    Use col_date(format = "%Y-%d-%m") within cols() as the col_types argument of read_csv().

In "desserts.csv", the variable uk_airdate is formatted like "17 August 2010". Let's parse, then cast this variable!

Later on in Chapter 4, we'll showcase some functions from the lubridate package to help you extract data from dates, once they are cast properly.
Instructions 1/3
30 XP

    1
    2
    3

Use ?parse_date to find the right format to parse the date "17 August 2010".


```{r}
library(readr)
# Find format to parse "17 August 2010" 
fecha <- parse_date("17 August 2010", format = "%d %B %Y")
fecha
class(fecha)

```

Leer data set desserts

```{r}
library(readr)

# Create bakeoff but skip first row
desserts <- read_csv("desserts_tidy.csv")


```

    
Exercise
Recode a character variable

In this exercise, you'll recode() the nut variable in the desserts data. This is a character variable that tells us, for each bake, whether a nut was a key ingredient and if so, what kind of nut!

Remember that the arguments of recode() are the variable that you want to recode, and then an expression of the form old_value = new_value. recode() is most useful when used inside a mutate() to create a new variable or reassign the old one.

Often you want to recode a certain value as missing. In this case, it is important to preserve the data type of the variable. R has built-in words to indicate missing data of various types, for example NA_character_ for missing value for a character variable, or NA_integer_ for missing integer data. The NA value is of the logical type.

Remember that you can use ?function_name_here to read more about arguments and how to use them.
Instructions 1/3
30 XP

    1
    2
    3

    Count the distinct values in the nut variable, sorting the result based on the number of rows. How many bakes included filberts as the nut?
```{r}
library(dplyr)
# Count rows grouping by nut variable
desserts %>% count(nut) %>% arrange(desc(n))
```
Recode a character variable

In this exercise, you'll recode() the nut variable in the desserts data. This is a character variable that tells us, for each bake, whether a nut was a key ingredient and if so, what kind of nut!

Remember that the arguments of recode() are the variable that you want to recode, and then an expression of the form old_value = new_value. recode() is most useful when used inside a mutate() to create a new variable or reassign the old one.

Often you want to recode a certain value as missing. In this case, it is important to preserve the data type of the variable. R has built-in words to indicate missing data of various types, for example NA_character_ for missing value for a character variable, or NA_integer_ for missing integer data. The NA value is of the logical type.

Remember that you can use ?function_name_here to read more about arguments and how to use them.
Instructions 2/3
21 XP

    2
    3

    What is a "filbert", you may ask? Recode "filbert" to "hazelnut", store the result in the same nut column, and count again.
```{r}
# Count rows grouping by nut variable
desserts %>% 
    count(nut, sort = TRUE)
    
# Recode filberts as hazelnuts
desserts_2 <- desserts %>% 
  mutate(nut = recode(nut, "filbert" = "hazelnut"))

# Count rows again 
desserts_2 %>% 
    count(nut, sort = TRUE)
```
Edit your code to also recode "no nut" as a missing character value, and count again to check your work.

```{r}
# Count rows grouping by nut variable
desserts %>% 
    count(nut, sort = TRUE)
    
# Edit code to recode "no nut" as missing
desserts_2 <- desserts %>% 
  mutate(nut = recode(nut, "filbert" = "hazelnut", 
                           "no nut" = NA_character_))

# Count rows again 
desserts_2 %>% 
    count(nut, sort = TRUE)
```

