---
title: "0103 ImportExcelData"
author: "Renato Erazo"
date: "16/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Install packge

```{r}
install.packages("readxls")
libray(readxl)
```


List the sheets of an Excel file
Before you can start importing from Excel, you should find out which sheets are available in the workbook. You can use the excel_sheets() function for this.

You will find the Excel file urbanpop.xlsx in your working directory (type dir() to see it). This dataset contains urban population metrics for practically all countries in the world throughout time (Source: Gapminder). It contains three sheets for three different time periods. In each sheet, the first row contains the column names.

Load the readxl package using library(). It's already installed on DataCamp's servers.
Use excel_sheets() to print out the names of the sheets in urbanpop.xlsx.

```{r}
# Load the readxl package
library(readxl)

# Print the names of all worksheets
excel_sheets("urbanpop.xlsx")
```

Import an Excel sheet
Now that you know the names of the sheets in the Excel file you want to import, it is time to import those sheets into R. You can do this with the read_excel() function. Have a look at this recipe:

data <- read_excel("data.xlsx", sheet = "my_sheet")
This call simply imports the sheet with the name "my_sheet" from the "data.xlsx" file. You can also pass a number to the sheet argument; this will cause read_excel() to import the sheet with the given sheet number. sheet = 1 will import the first sheet, sheet = 2 will import the second sheet, and so on.

In this exercise, you'll continue working with the urbanpop.xlsx file.

The code to import the first and second sheets is already included. Can you add a command to also import the third sheet, and store the resulting data frame in pop_3?
Store the data frames pop_1, pop_2 and pop_3 in a list, that you call pop_list.
Display the structure of pop_list.

```{r}
# The readxl package is already loaded

# Read the sheets, one by one
pop_1 <- read_excel("urbanpop.xlsx", sheet = 1)
pop_2 <- read_excel("urbanpop.xlsx", sheet = 2)
pop_3 <- read_excel("urbanpop.xlsx", sheet = 3)


# Put pop_1, pop_2 and pop_3 in a list: pop_list
pop_list <- list(pop_1,pop_2,pop_3)

# Display the structure of pop_list
str(pop_list)
```

## lapply and read

The read_excel() function is called multiple times on the "data.xlsx" file and each sheet is loaded in one after the other. The result is a list of data frames, each data frame representing one of the sheets in data.xlsx.

You're still working with the urbanpop.xlsx file.

Use lapply() in combination with excel_sheets() and read_excel() to read all the Excel sheets in "urbanpop.xlsx". Name the resulting list pop_list.
Print the structure of pop_list.

```{r}
# The readxl package is already loaded

# Read all Excel sheets with lapply(): pop_list

pop_list <- lapply(excel_sheets("urbanpop.xlsx"), read_excel,path = "urbanpop.xlsx")

# Display the structure of pop_list
str(pop_list)
```
The col_names argument
Apart from path and sheet, there are several other arguments you can specify in read_excel(). One of these arguments is called col_names.

By default it is TRUE, denoting whether the first row in the Excel sheets contains the column names. If this is not the case, you can set col_names to FALSE. In this case, R will choose column names for you. You can also choose to set col_names to a character vector with names for each column. It works exactly the same as in the readr package.

You'll be working with the urbanpop_nonames.xlsx file. It contains the same data as urbanpop.xlsx but has no column names in the first row of the excel sheets.

#Exercise

Import the first Excel sheet of "urbanpop_nonames.xlsx" and store the result in pop_a. Have R set the column names of the resulting data frame itself.
Import the first Excel sheet of urbanpop_nonames.xlsx; this time, use the cols vector that has already been preparedfor you to specify the column names. Store the resulting data frame in pop_b.
Print out the summary of pop_a.
Print out the summary of pop_b. Can you spot the difference with the other summary?




```{r}

# The readxl package is already loaded

# Import the first Excel sheet of urbanpop_nonames.xlsx (R gives names): pop_a
pop_a <- read_excel("urbanpop.xlsx", col_names = FALSE)

# Import the first Excel sheet of urbanpop_nonames.xlsx (specify col_names): pop_b
cols <- c("country", paste0("year_", 1960:1966))
pop_b <- read_excel("urbanpop.xlsx", col_names = cols)

# Print the summary of pop_a
summary(pop_a)

# Print the summary of pop_b
summary(pop_b)

```

The skip argument
Another argument that can be very useful when reading in Excel files that are less tidy, is skip. With skip, you can tell R to ignore a specified number of rows inside the Excel sheets you're trying to pull data from. Have a look at this example:

read_excel("data.xlsx", skip = 15)
In this case, the first 15 rows in the first sheet of "data.xlsx" are ignored.

If the first row of this sheet contained the column names, this information will also be ignored by readxl. Make sure to set col_names to FALSE or manually specify column names in this case!

The file urbanpop.xlsx is available in your directory; it has column names in the first rows.

Import the second sheet of "urbanpop.xlsx", but skip the first 21 rows. Make sure to set col_names = FALSE. Store the resulting data frame in a variable urbanpop_sel.
Select the first observation from urbanpop_sel and print it out.


```{r}
# The readxl package is already loaded

# Import the second sheet of urbanpop.xlsx, skipping the first 21 rows: urbanpop_sel

urbanpop_sel <- read_excel("urbanpop.xlsx",sheet = 2, skip = 21)

# Print out the first observation from urbanpop_sel
urbanpop_sel[1,]
```

```{r}
# The readxl package is already loaded

# Import the second sheet of urbanpop.xlsx, skipping the first 21 rows: urbanpop_sel

urbanpop_sel <- read_excel("urbanpop.xlsx",sheet = 2, col_names = FALSE,skip = 21)

# Print out the first observation from urbanpop_sel
urbanpop_sel[1,]
```



## ========= Standar package for import excel

```{r}
install.packages("gdata")
library(gdata)

```
Import a local file
In this part of the chapter you'll learn how to import .xls files using the gdata package. Similar to the readxl package, you can import single Excel sheets from Excel sheets to start your analysis in R.

You'll be working with the urbanpop.xls dataset, the .xls version of the Excel file you've been working with before. It's available in your current working directory.

Instructions
100 XP
Load the gdata package with library(). gdata and Perl are already installed on DataCamp's Servers.
Import the second sheet, named "1967-1974", of "urbanpop.xls" with read.xls(). Store the resulting data frame as urban_pop.
Print the first 11 observations of urban_pop with head().

```{r}
# Load the gdata package
library(gdata)

# Import the second sheet of urbanpop.xls: urban_pop
urban_pop <- read.xls("urbanpop.xls",sheet = 2)

# Print the first 11 observations using head()

head(urban_pop)
```

```{r}
# Column names for urban_pop
columns <- c("country", paste0("year_", 1967:1974))

# Finish the read.xls call
urban_pop <- read.xls("urbanpop.xls", sheet = 2,
                      skip = 50, header = FALSE, stringsAsFactors = FALSE,
                      col.names = columns)

# Print first 10 observation of urban_pop
head(urban_pop)
```

